\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-07 paper.qmd
Updated Five Gau second paragraph

In this section, the evaluation focuses on how effectively the tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure. The data consists of five $4\text{-}D$ clusters, each with $1000$ points generated from a multivariate normal distribution with a small variance of 0.0025. The first cluster is centered at $(0, 0, 0, 0)$, the second at $(1, 0, 0, 0)$, the third at $(0, 1, 0, 0)$, the fourth at $(0, 0, 1, 0)$, and the fifth at $(0, 0, 0, 1)$.

@@

In the t-SNE 2D layout, the distances between clusters appear smaller, leading to more tightly packed clusters, which is typical of t-SNE's focus on preserving local structures. This often results in less emphasis on the global distances between clusters. On the other hand, UMAP tends to maintain larger separations between clusters, better reflecting the global structure and preserving the relative distances between clusters more effectively. PaCMAP, meanwhile, balances these aspects by maintaining the distance between clusters and positioning them correctly, effectively capturing both global and local structures.

@@

```{r}
#| label: fig-five-gau-nldr-layouts
#| fig-height: 5
#| fig-width: 15
#| fig-pos: H
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"

tsne_gau + umap_gau + pacmap_gau +
  plot_layout(ncol = 3)
```

```{r}
#| label: fig-five-gau-projs
#| fig-height: 10
#| fig-width: 15
#| fig-pos: H
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"

five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + five_gau_proj_tsne_model2 + five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
  plot_layout(guides = "collect", nrow = 2) &
  theme(legend.position='none')
```

>>>>

In this section, the evaluation focuses on how effectively the tSNE, UMAP, and PaCMAP algorithms preserve the clustering structure. The data consists of five $4\text{-}D$ clusters, each with $1000$ points generated from a multivariate normal distribution with a small variance of $0.0025$. The first cluster is centered at $(0, 0, 0, 0)$, the second at $(1, 0, 0, 0)$, the third at $(0, 1, 0, 0)$, the fourth at $(0, 0, 1, 0)$, and the fifth at $(0, 0, 0, 1)$.

@@

In the tSNE layout (@fig-five-gau-projs (a1)), the distances between clusters appear smaller, leading to more tightly packed clusters, which is typical of tSNE's focus on preserving local structures. This often results in less emphasis on the global distances between clusters. On the other hand, UMAP (@fig-five-gau-projs (b1)) tends to maintain larger separations between clusters, better reflecting the global structure and preserving the relative distances between clusters more effectively. In contrast, PaCMAP (@fig-five-gau-projs (c1)) aim to balance these aspects by maintaining the distance between clusters and accurately positioning them, effectively capturing both global and local structures.

@@

```{r}
#| label: fig-five-gau-projs
#| fig-height: 15
#| fig-width: 15
#| fig-pos: H
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"

tsne_gau + umap_gau + pacmap_gau +
five_gau_proj_tsne_model1 + five_gau_proj_umap_model1 + five_gau_proj_pacmap_model1 + five_gau_proj_tsne_model2 + five_gau_proj_umap_model2 + five_gau_proj_pacmap_model2 +
  plot_layout(guides = "collect", nrow = 3) &
  theme(legend.position='none')
```

### 2024-09-07 paper.tex
Updated Five Gau second paragraph

In this section, the evaluation focuses on how effectively the tSNE,
UMAP, and PaCMAP algorithms preserve the clustering structure. The data
consists of five \(4\text{-}D\) clusters, each with \(1000\) points
generated from a multivariate normal distribution with a small variance
of 0.0025. The first cluster is centered at \((0, 0, 0, 0)\), the second
at \((1, 0, 0, 0)\), the third at \((0, 1, 0, 0)\), the fourth at
\((0, 0, 1, 0)\), and the fifth at \((0, 0, 0, 1)\).

In the t-SNE 2D layout, the distances between clusters appear smaller,
leading to more tightly packed clusters, which is typical of t-SNE's
focus on preserving local structures. This often results in less
emphasis on the global distances between clusters. On the other hand,
UMAP tends to maintain larger separations between clusters, better
reflecting the global structure and preserving the relative distances
between clusters more effectively. PaCMAP, meanwhile, balances these
aspects by maintaining the distance between clusters and positioning
them correctly, effectively capturing both global and local structures.

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-five-gau-nldr-layouts-1.pdf}

\caption{\label{fig-five-gau-nldr-layouts}(a) tSNE, (b) UMAP, and (c)
PaCMAP}

>>>>

In this section, the evaluation focuses on how effectively the tSNE,
UMAP, and PaCMAP algorithms preserve the clustering structure. The data
consists of five \(4\text{-}D\) clusters, each with \(1000\) points
generated from a multivariate normal distribution with a small variance
of \(0.0025\). The first cluster is centered at \((0, 0, 0, 0)\), the
second at \((1, 0, 0, 0)\), the third at \((0, 1, 0, 0)\), the fourth at
\((0, 0, 1, 0)\), and the fifth at \((0, 0, 0, 1)\).

In the tSNE layout (Figure~\ref{fig-five-gau-projs} (a1)), the distances
between clusters appear smaller, leading to more tightly packed
clusters, which is typical of tSNE's focus on preserving local
structures. This often results in less emphasis on the global distances
between clusters. On the other hand, UMAP
(Figure~\ref{fig-five-gau-projs} (b1)) tends to maintain larger
separations between clusters, better reflecting the global structure and
preserving the relative distances between clusters more effectively. In
contrast, PaCMAP (Figure~\ref{fig-five-gau-projs} (c1)) aim to balance
these aspects by maintaining the distance between clusters and
accurately positioning them, effectively capturing both global and local
structures.

@@
